Supplementary Materials for Differences in the Risk of Grade Retention for Biracial and Monoracial Students in the US, 2010-2019

### 2023-01-09 products/paper/main.qmd
Small changes for Soc Sci submission

---
title: Differences in the Risk of Grade Retention for Biracial and Monoracial Students in the US, 2010-2019
author: 
  - Aaron Gullickson, *University of Oregon*
abstract: Understanding how outcomes for biracial individuals compare to their monoracial peers is critical for understanding how patterns of racial inequality in the contemporary United States might be shifting. Yet, we know very little about the life chances of biracial individuals due to limitations in most available data sources. In this article, I utilize American Community Survey data from 2010-2019 to examine the risk of being clearly behind expected grade among biracial and monoracial K-12 students, helping to fill a gap in our understanding. With large sample sizes for most biracial groups, I am able to estimate grade retention risk for biracial students with enough precision to differentiate even modest differences in risk relative to monoracial groups. The results indicate that for most biracial groups, biracial students have risk similar to their monoracial constituent group with lower risk. Although biracial students tend to have favorable family resource characteristics, controlling for these characteristics does little to change the overall placement of their outcomes.
thanks: The corresponding author can be reached at aarong@uoregon.edu. Supplementary materials provided with this article include full model results upon which figures are based as well as any sensitivity analysis described herein. All code and data for this project are available at [https://osf.io/4fevt/?view_only=4abc6d86595c4313a8d4792471e9bc0d](https://osf.io/4fevt/?view_only=4abc6d86595c4313a8d4792471e9bc0d).
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    fig-height: 4
    fig-pos: t
    keep-tex: false
  docx:
    toc: false
bibliography: ../project.bib
execute: 
  eval: true
  echo: false
  warning: false
  error: false
prefer-html: true
editor: source
csl: american-sociological-association.csl
---

@@

```{r monoracial-results}
#| results: asis
#| label: tbl-monoracial
#| tbl-cap: Average marginal differences in the probability of being clearly behind expected grade by monoracial group, relative to a monoracial white student
texreg(map(results_monoracial[-1], convert_marg),
#wordreg(map(results_monoracial[-1], convert_marg),
       digits=3,
       caption = "Average marginal differences in the probability of being clearly behind expected grade by monoracial group, relative to a monoracial white student",
       caption.above=TRUE,
       label="tbl-monoracial",  #TODO: This does not work
       float.pos="t",
#       file="regtable.doc",
       custom.gof.rows=
         list("State fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Location fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Grade fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Year linear effects" = c("Yes","Yes","Yes","Yes"),
              "Year x grade effects"  = c("Yes","Yes","Yes","Yes"),
              "Nativity and language" = c("No","Yes","Yes","Yes"),
              "Parent's education" = c("No","No","Yes","Yes"),
              "Other family resources" = c("No","No","No","Yes")))
```

>>>>

---
title: Differences in the Risk of Grade Retention for Biracial and Monoracial Students in the US, 2010-2019
#author: 
#  - Aaron Gullickson, *University of Oregon*
abstract: Understanding how outcomes for biracial individuals compare to their monoracial peers is critical for understanding how patterns of racial inequality in the contemporary United States might be shifting. Yet, we know very little about the life chances of biracial individuals due to limitations in most available data sources. In this article, I utilize American Community Survey data from 2010-2019 to examine the risk of being clearly behind expected grade among biracial and monoracial K-12 students, helping to fill a gap in our understanding. With large sample sizes for most biracial groups, I am able to estimate grade retention risk for biracial students with enough precision to differentiate even modest differences in risk relative to monoracial groups. The results indicate that for most biracial groups, biracial students have risk similar to their monoracial constituent group with lower risk. Although biracial students tend to have favorable family resource characteristics, controlling for these characteristics does little to change the overall placement of their outcomes.
thanks: Supplementary materials provided with this article include full model results upon which figures are based as well as any sensitivity analysis described herein. All code and data for this project are available at [https://osf.io/4fevt/?view_only=4abc6d86595c4313a8d4792471e9bc0d](https://osf.io/4fevt/?view_only=4abc6d86595c4313a8d4792471e9bc0d).
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    fig-height: 4
    fig-pos: t
    linestretch: 1.5
    keep-tex: false
  docx:
    toc: false
  html:
    toc: false
bibliography: ../project.bib
execute: 
  eval: true
  echo: false
  warning: false
  error: false
prefer-html: true
editor: source
csl: american-sociological-association.csl
---

@@

```{r monoracial-results}
#| results: asis
#| label: tbl-monoracial
#| tbl-cap: Average marginal differences in the probability of being clearly behind expected grade by monoracial group, relative to a monoracial white student
texreg(map(results_monoracial[-1], convert_marg),
#htmlreg(map(results_monoracial[-1], convert_marg),
#wordreg(map(results_monoracial[-1], convert_marg),
       digits=3,
       caption = "Average marginal differences in the probability of being clearly behind expected grade by monoracial group, relative to a monoracial white student",
       caption.above=TRUE,
       label="tbl-monoracial",  #TODO: This does not work
       float.pos="t",
#       file="regtable.doc",
       custom.gof.rows=
         list("State fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Location fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Grade fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Year linear effects" = c("Yes","Yes","Yes","Yes"),
              "Year x grade effects"  = c("Yes","Yes","Yes","Yes"),
              "Nativity and language" = c("No","Yes","Yes","Yes"),
              "Parent's education" = c("No","No","Yes","Yes"),
              "Other family resources" = c("No","No","No","Yes")))
```

### 2023-01-09 products/paper/supp_materials.qmd
Small changes for Soc Sci submission

```{r}
#| results: asis
#| label: tbl-full-models
#| tbl-cap: Log-odds ratios from full models predicting a student being clearly behind expected grade. All models used sample weights and adjust for design effects from sample weight variance and clustering of students within the same household.
wordreg(lapply(list(model0_summary, model1_summary, model2_summary, 
                   model3_summary, model4_summary), convert_model),
       digits = 3,
       longtable = TRUE,
       use.packages = FALSE,
       caption.above = TRUE,
       file="full_reg.doc",
       custom.coef.names = 
         c("Intercept","Black","Indigenous","Asian","Latino","Black/White",
           "Black/Indigenous","Black/Latino", "Black/Asian",
           "White/Indigenous","White/Latino","White/Asian",
           "Indigenous/Latino","Indigenous/Asian","Latino/Asian",
           "Year (2010 origin)", "1st grade","2nd grade","3rd grade",
           "4th grade","5th grade","6th grade","7th grade","8th grade",
           "9th grade","10th grade","11th grade","12th grade","Alaska",
           "Arizona","Arkansas","California","Colorado","Connecticut",
           "Washington DC","Delaware","Florida","Georgia","Hawaii","Idaho",
           "Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana",
           "Maine","Maryland","Massachusetts","Michigan","Minnesota",
           "Mississippi","Missouri","Montana","Nebraska","Nevada",
           "New Hampshire","New Jersey","New Mexico","New York",
           "North Carolina","North Dakota","Ohio","Oklahoma","Oregon",
           "Pennsylvania","Rhode Island","South Carolina","South Dakota",
           "Tennessee","Texas","Utah","Vermont","Virginia","Washington",
           "West Virginia","Wisconsin","Wyoming","Non-metro","Central city",
           "Metro non-central","Metro indeterminable","Female",
           "Year x 1st grade","Year x 2nd grade","Year x 3rd grade",
           "Year x 4th grade","Year x 5th grade","Year x 6th grade",
           "Year x 7th grade","Year x 8th grade","Year x 9th grade",
           "Year x 10th grade","Year x 11th grade","Year x 12th grade",
           "Foreign-born","Mother foreign-born","Father foreign-born",
           "Speaks English somewhat", "Speaks English well", 
           "Mother speaks somewhat", "Mother speaks well",
           "Father speaks somewhat", "Father speaks well",
           "Mother High school diploma", "Mother AA degree", 
           "Mother BA degree","Mother Grad degree",
           "Father High school diploma", "Father AA degree", 
           "Father BA degree","Father Grad degree",
           "Family income (10K USD, sq. root)", "Owns home",
           "parents married"),
       groups = list("Race (ref. White)"=2:15, 
                     "Grade (ref. Kindergarten)"=17:28,
                     "State (ref. Alabama)"=29:78,
                     "Location (ref. Indeterminable)"=79:82,
                     "English proficiency (ref. none)"=99:100,
                     "Mother's English proficiency (ref. none)"=101:102,
                     "Father's English proficiency (ref. none)"=103:104,
                     "Mother's highest degree (ref. none)"=105:108,
                     "Father's highest degree (ref. none)"=109:112))
```

>>>>

```{r}
#| results: asis
#| label: tbl-full-models
#| tbl-cap: Log-odds ratios from full models predicting a student being clearly behind expected grade. All models used sample weights and adjust for design effects from sample weight variance and clustering of students within the same household.
texreg(lapply(list(model0_summary, model1_summary, model2_summary, 
                   model3_summary, model4_summary), convert_model),
       digits = 3,
       longtable = TRUE,
       use.packages = FALSE,
       caption.above = TRUE,
       #file="full_reg.doc",
       custom.coef.names = 
         c("Intercept","Black","Indigenous","Asian","Latino","Black/White",
           "Black/Indigenous","Black/Latino", "Black/Asian",
           "White/Indigenous","White/Latino","White/Asian",
           "Indigenous/Latino","Indigenous/Asian","Latino/Asian",
           "Year (2010 origin)", "1st grade","2nd grade","3rd grade",
           "4th grade","5th grade","6th grade","7th grade","8th grade",
           "9th grade","10th grade","11th grade","12th grade","Alaska",
           "Arizona","Arkansas","California","Colorado","Connecticut",
           "Washington DC","Delaware","Florida","Georgia","Hawaii","Idaho",
           "Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana",
           "Maine","Maryland","Massachusetts","Michigan","Minnesota",
           "Mississippi","Missouri","Montana","Nebraska","Nevada",
           "New Hampshire","New Jersey","New Mexico","New York",
           "North Carolina","North Dakota","Ohio","Oklahoma","Oregon",
           "Pennsylvania","Rhode Island","South Carolina","South Dakota",
           "Tennessee","Texas","Utah","Vermont","Virginia","Washington",
           "West Virginia","Wisconsin","Wyoming","Non-metro","Central city",
           "Metro non-central","Metro indeterminable","Female",
           "Year x 1st grade","Year x 2nd grade","Year x 3rd grade",
           "Year x 4th grade","Year x 5th grade","Year x 6th grade",
           "Year x 7th grade","Year x 8th grade","Year x 9th grade",
           "Year x 10th grade","Year x 11th grade","Year x 12th grade",
           "Foreign-born","Mother foreign-born","Father foreign-born",
           "Speaks English somewhat", "Speaks English well", 
           "Mother speaks somewhat", "Mother speaks well",
           "Father speaks somewhat", "Father speaks well",
           "Mother High school diploma", "Mother AA degree", 
           "Mother BA degree","Mother Grad degree",
           "Father High school diploma", "Father AA degree", 
           "Father BA degree","Father Grad degree",
           "Family income (10K USD, sq. root)", "Owns home",
           "parents married"),
       groups = list("Race (ref. White)"=2:15, 
                     "Grade (ref. Kindergarten)"=17:28,
                     "State (ref. Alabama)"=29:78,
                     "Location (ref. Indeterminable)"=79:82,
                     "English proficiency (ref. none)"=99:100,
                     "Mother's English proficiency (ref. none)"=101:102,
                     "Father's English proficiency (ref. none)"=103:104,
                     "Mother's highest degree (ref. none)"=105:108,
                     "Father's highest degree (ref. none)"=109:112))
```
